---
title: 'CB Geomarket'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include = F}
knitr::opts_chunk$set(message = F, warning = F)

options(scipen = 10000)
```

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(haven)
library(labelled)
library(tidycensus)
library(sf)
library(tigris)
library(maptools)
```


# Load data

```{r}
data_dir <- file.path('..', '..', 'data')
eps_data_dir <- file.path(data_dir, 'eps_market')
```

```{r}
# Geomarket to zip conversion
eps_zip <- read_excel(path = file.path(eps_data_dir, 'eps.xls'), sheet = 'Sheet 1') %>% 
  mutate(eps = if_else(str_sub(eps, 3, 3) == '0', str_c(str_sub(eps, 1, 2), ' ', str_sub(eps, 4, 4)), eps))

eps_zip
```

```{r}
# Zip code shapes (2010)
zcta_2010 <- zctas(
  cb = T, 
  year = 2010
)

zcta_2010
```

```{r}
# Zip code income data (2015)
income_2015_zcta <- get_acs(
  geography = 'zcta',
  variables = 'B19013_001',
  year = 2015,
  geometry = F,
  survey = 'acs5'
)

income_2015_zcta
```

```{r}
# Aggregate income data to geomarket level
income_2015_eps <- eps_zip %>% 
  inner_join(income_2015_zcta, by = c('zip' = 'GEOID')) %>%
  group_by(eps) %>% 
  summarize(
    mean_inc = mean(estimate, na.rm = T)
  ) %>% 
  ungroup()

income_2015_eps
income_2015_eps %>% arrange(desc(mean_inc)) %>% print(n=100)
```

```{r}
# CA data
eps_zip_CA <- zcta_2010 %>%  # has to be left join (right/inner loses attributes on geometry)
  left_join(eps_zip, by = c('ZCTA5' = 'zip')) %>% 
  filter(str_detect(eps, '^CA'))

eps_zip_CA
```

```{r}
myzip <- readShapePoly(file.path(eps_data_dir, 'US_Zip_-_Lower_48.shp'))

myzip@data <- data.frame(myzip@data, eps_zip[match(as.character(myzip@data$ZIP), eps_zip$zip),])
myzip@data
myzip_CA <- myzip[!is.na(myzip@data$eps) & str_detect(myzip@data$eps, '^CA'), ]
```


# Plot CA (zip-code level)

## Data from `US_Zip_-_Lower_48.shp`

Matches zip codes from [CA state geoportal](https://gis.data.ca.gov/datasets/CDEGIS::california-zip-codes/explore?location=38.510705%2C-120.973098%2C6.71):

```{r}
plot(myzip_CA, col = 'pink')
```

## Data from Census Bureau

Matches zip codes from [Census Bureau](https://hub.arcgis.com/datasets/7f3aa1bd8d1f4915b928ff4da5f3f3bb/explore?location=38.818303%2C-119.108380%2C6.90):

```{r}
eps_zip_CA %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

Color by zip code-level income:

```{r}
eps_zip_CA %>% 
  left_join(income_2015_zcta, by = c('NAME' = 'GEOID')) %>% 
  ggplot(aes(fill = estimate)) +
  geom_sf(color = NA) +
  scale_fill_gradient(low = 'lightyellow', high = 'red')
```


# Plot CA (geomarket level)

No difference between aggregate functions:

```{r}
# Aggregate by identity function
eps_shapes <- eps_zip_CA %>%
  aggregate(by = list(eps_zip_CA$eps), FUN = identity) %>% 
  select(-eps) %>% rename(eps = Group.1)

eps_shapes %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

```{r}
# Aggregate by mean function
eps_shapes <- eps_zip_CA %>%
  aggregate(by = list(eps_zip_CA$eps), FUN = mean) %>% 
  select(-eps) %>% rename(eps = Group.1)

eps_shapes %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

Color by geomarket-level income:

```{r}
eps_shapes %>% 
  left_join(income_2015_eps, by = 'eps') %>% 
  ggplot(aes(fill = mean_inc)) +
  geom_sf(color = NA) +
  geom_sf_label(aes(label = eps)) +
  scale_fill_gradient(low = 'lightyellow', high = 'red')
```

Color by zip code-level income w/ geomarket outline (i.e., within geomarket):

```{r}
eps_zip_CA %>% 
    left_join(income_2015_zcta, by = c('NAME' = 'GEOID')) %>% 
    ggplot() +
    geom_sf(aes(fill = estimate), color = NA) +
    geom_sf(data = eps_shapes, fill = NA, color = 'black') +
    scale_fill_gradient(low = 'lightyellow', high = 'red')
```
